---
title: "Venezia High Waters EDA"
author: "Luis Bronchal"
output:
  html_document: default
  html_notebook: default
  pdf_document: default
---

### Summary 

This is a first Exploratory Data Analisys for the aggregated data of Venezia High Waters.
Original data come from [Centro Previsioni e Segnalazioni Maree](http://www.comune.venezia.it/archivio/25419)

### Checking the data 

```{r, message=FALSE, warning=FALSE}
library(TSA)
library(xts)
library(lubridate)
library(ggplot2)
library(forecast)
library(readr)
library(dplyr)
library(tidyr)
```

We load the data and check the its quality

```{r}
data <- read_csv(file = "../output/venezia.csv",
                 col_types = cols_only(col_datetime(), col_number()))

data_xts <- xts(order.by=data$datetime,
                      x=data$level)
attr(data_xts, 'frequency') <- 24
```

```{r}
first(data_xts)
end(data_xts)
```

There are not any NA values:
```{r}
anyNA(data_xts)
```

We look if there are duplicated:

```{r}
dup <- duplicated(index(data_xts), fromLast=TRUE)
sum(dup)
```

We look what the duplicated values are:

```{r}
data_xts[index(data_xts[dup,])]
```

We conserve only the first time a duplicated register was taken:
```{r}
data_xts <- data_xts[!dup]
```

We look for holes in the data
```{r}
data_xts_index <- index(data_xts)
all_points <- seq(first(data_xts_index), last(data_xts_index), by="hour") 

# it looks like there are holes in the data
length(all_points) == length(index(data_xts))

# we complete the holes with the previous observation value
venezia_xts <- xts(order.by=all_points)
venezia_xts <- merge(data_xts, venezia_xts, join = "right")
venezia_xts <- na.locf(venezia_xts)
length(all_points) == length(index(venezia_xts))
```

### Exploratory Data Analysis

The historical minimum and maximum level:
```{r}
range(venezia_xts)
```

The mean level:
```{r}
mean(venezia_xts)
```

We take a look to some graphs
```{r}
boxplot(coredata(venezia_xts))
```

To plot all the data is confusing
```{r}
ts.plot(venezia_xts)
```

Let's see one month:
```{r}
ts.plot(venezia_xts["2015-01"])
```


```{r}


# compute the Fourier Transform
p <- periodogram(venezia_xts)
dd <- data.frame(freq=p$freq, spec=p$spec, time=1/p$freq)
order <- dd[order(-dd$spec),]
top10 <- head(order, 10)
top10

# references:
# https://anomaly.io/detect-seasonality-using-fourier-transform-r/
# http://rpubs.com/sinhrks/plot_tsstats
#

```

The main frequencies are around each 12 and 24 hours.


```{r}
data2 <- as.data.frame(venezia_xts)
data2$datetime <- rownames(data2)
rownames(data2) <- NULL
names(data2) <- c("level", "datetime")

data2$year <- year(data2$datetime)
data2$month <- month(data2$datetime)
data2$day <- day(data2$datetime)
data2$weekday <- factor(wday(data2$datetime), levels=1:7, labels=c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday" , "Saturday", "Sunday"))
data2$hour <- hour(data2$datetime)
data2 <- data2[c("datetime", "year", "month", "day", "weekday", "level")]
```

We remove 2016 data because it has only 1 observation (first hour of the first day)
```{r}
data2 <- data2 %>% filter(year<2016)
```

We see the variation of the water levels through years. It looks like there are some kind of oscilation where some years have higher levels 
```{r}
# per year
ggplot(data2, aes(x=year, y=level, group=year, col=factor(year))) + geom_boxplot() 
```

If we look it monthly, it looks like the months with low level are January, February and March
```{r}
# per month
ggplot(data2, aes(x=as.factor(month), y=level, group=month, col=as.factor(month))) + geom_boxplot() + scale_color_discrete("month", labels=month.abb) +
  scale_x_discrete("month", labels=month.abb)
```

It looks like there are not great variations if we consider the day of the week:

```{r}
# per weekday
ggplot(data2, aes(x=weekday, y=level, group=weekday, col=weekday)) + geom_boxplot()
```

```{r}
data_yearly <- data2 %>% group_by(year) %>% 
  summarise(max=max(level), min=min(level), mean=mean(level)) %>% 
  gather(key, value, -year)
```

It looks like there is an increase in the mean level year after year:
```{r}
data_yearly %>% ggplot(aes(x=year, y=value, fill=as.factor(key))) + geom_bar(stat="identity") +
  scale_fill_discrete("Water level") + scale_y_continuous("level")
```

If we calculate linear regresion it looks like there is an increase in the max and min level also year after year.
```{r}
data_yearly %>% ggplot(aes(x=year, y=value, col=as.factor(key))) + geom_point() + 
  geom_smooth(method="lm") + scale_color_discrete("Water level") + 
  scale_y_continuous("level")
```

